---
title: "Homework 2"
author: "Johnson ODEJIDE"
date: "2023-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
library(MASS)
```
## 1

```{r Solution Concentration}
y <-  c(0.07, 0.09, 0.08, 0.16, 0.17, 0.21, 0.49, 0.58, 0.53, 1.22, 1.15, 1.07, 2.84, 2.57, 3.10)

x <- c(9.0, 9.0, 9.0, 7.0, 7.0, 7.0, 5.0, 5.0, 5.0, 3.0, 3.0, 3.0, 1.0, 1.0, 1.0)

# length(y) - Verify the length of vector y - 15

# length(x) - Verify the length of vector x - 15
```

```{r}
# Fit a linear regression function
lm.fit <- lm(y ~ x)
summary(lm.fit)
```
**The Linear Regression Function is given as:**

`y(hat) = 2.5753 - 0.3240x`

#### F-Test 

```{r}
anova(lm.fit)
```

```{r}
f.value <- 55.994

f.critical <- qf(p = 0.025, df1 = 1, df2 = 13, lower.tail = FALSE)

print(paste("F value = ", f.value, "F Critical = ", f.critical))

ifelse(f.value > f.critical, "Reject the null hypothesis", "Fail to reject the null hypothesis")
```


## 2

```{r Sales Growth}
X <- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9)

Y <- c(98, 135, 162, 178, 221, 232, 283, 300, 374, 395)
``` 

### a.

```{r}
sales_growth <- data.frame(X, Y)

# Scatter plot
sales_growth %>% 
  ggplot(aes(x = X, y = Y)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  theme_bw()

# scatter.smooth(Y~X)
```
_There appears to be a positive relationship but the linear relation doesn't seem to be adequate_

### b

#### Box-Cox

```{r "Box-Cox for Sales Growth"}
box_cox <- boxcox(Y ~ X)

lambda <- box_cox$x[which.max(box_cox$y)] 
lambda

new_model <- lm(((Y^lambda-1)/lambda) ~ X) 
new_model

```
_The appropriate power transformation of Y is 0.5050_

#### SSE  for Lambdas

```{r}
lambda.3 <- (lm(((Y^.3 - 1) / .3) ~ X))
anova(lambda.3) # 0.987

lambda.4 <- lm(((Y^.4 - 1) / .4) ~ X) 
summary(lambda.4) # 0.9885

lambda.5 <- lm(((Y^.5 - 1) / .5) ~ X) 
summary(lambda.5) # 0.9891

lambda.6 <- lm(((Y^.6 - 1) / .6) ~ X) 
summary(lambda.6) # 0.9889

lambda.7 <- lm(((Y^.7 - 1) / .7) ~ X) 
summary(lambda.7) # 0.9878

lambda <- c(.3, .4, .5, .6, .7)
SSE <- c(0.987, 0.9885, 0.9891, 0.9889, 0.9878)
data.frame(lambda, SSE)


```


_SSE = 0.987 when lamda = .3_

_SSE = 0.9885 when lamda = .4_

_SSE = 0.9891 when lamda = .5_

_SSE = 0.9889 when lamda = .6_

_SSE = 0.9878 when lamda = .7_

The suggested transformation of Y is when When lambda = .5, that is, sqrt(Y)

## c

```{r}
Y_trans <- lm(sqrt(Y) ~ X)
summary(Y_trans)
```

_log(Yhat) <- 10.261 + 1.076X_

```{r}
plot(Y_trans$residuals, sqrt(Y))
abline(0, 0, 'red')
```


## 3

```{r Matrices}
A <- matrix(c(1, 4,
              2, 6,
              3, 8), ncol = 2, byrow = TRUE)


B <- matrix(c(1, 3,
              1, 4,
              2, 5), ncol = 2, byrow = TRUE)


C <- matrix(c(3, 8, 1,
              5, 4, 0), ncol = 3, byrow = TRUE)
```

### i 

```{r}
res <- A + B
res
dim(res)
```
**3 by 2 matrix**

### ii

```{r}
res <- A - B
res
dim(res)
```
**3 by 2 matrix**

### iii

```{r}
res <- A %*% C
res
dim(res)
```
**3 by 3 matrix**

### iv

```{r}
res <- A %*% t(B)
res
dim(res)
```
**3 by 3 matrix**

### v

```{r}
res <- A %*% t(B)
res
dim(res)
```

**3 by 3 matrix**

### vi

```{r}
res <- t(B) %*% A
res
dim(res)
```

**2 by 2 matrix**

## 4

```{r Prediction and Confidence Intervals}
mtcars.fit <- lm(mpg ~ hp, data = mtcars)
mtcars.fit

newdf <- data.frame(hp = 62)
newdf

# Prediction Interval
predict(object = mtcars.fit, newdata = newdf, interval = "prediction") %>% 
  cbind(newdf)

# Confidence Interval
predict(object = mtcars.fit, newdata = newdf, interval = "confidence") %>% 
  cbind(newdf)
```
**Prediction Interval**
(17.6682, 34.06919)

**Confidence Interval**
(23.6308, 28.1066)
