---
title: "Homework 3"
author: "Johnson ODEJIDE"
date: "2023-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
library(tidyverse)
```

**1a)** Use matrix methods and r coding demonstrated in class to generate a linear regression model for the data.  What are the estimates bo and b1 ?

```{r}
Y <- matrix(c(16,
              5,
              10,
              15,
              13,
              22), ncol = 1, byrow = TRUE)

X <- matrix(c(1, 4, 
              1, 1,
              1, 2,
              1, 3,
              1, 3,
              1, 4), ncol = 2, byrow = TRUE)

t(X) -> transposeX
transposeX%*%X -> Product2
interceptandslope <- solve(Product2)%*%transposeX%*%Y

interceptandslope
```
Yhat = 0.4390 + 4.6098X

B0 = 0.4390, B1 = 4.6098

**1b.)** Use matrix methods and r coding to find the fitted values for your model.

```{r}
X %*% interceptandslope
```

**1c.)** Use matrix methods and r coding to find the residuals for your model.

```{r}
Y - X %*% interceptandslope
```

**1d)** Now use the standard lm r coding to confirm the results for bo and b1 that were obtained in part a)

```{r}
dt <- tribble(
  ~X,  ~Y,
  4,   16,
  1,    5,
  2,   10,
  3,   15,
  3,   13,
  4,   22
)

# dt

linear.model <- lm(Y ~ X, data = dt)
summary(linear.model)
```
Yhat = 0.4390 + 4.6098X

B0 = 0.4390, B1 = 4.6098

The values are the same as obtained in part (a)

```{r}
A <- matrix(c(5, 3, 1,
              2, 8, 5,
              4, 7, 2), 
            ncol = 3, byrow = TRUE)

B <- matrix(c(10, 4,
              6, 8,
              2, 1), ncol = 2, byrow = TRUE)


```

**2)** Use the matrices given above to confirm the property   (AB)T  =  BTAT

```{r}
AB_T <- t(A %*% B)
AB_T

B_T.A_T <- t(B) %*% t(A)
B_T.A_T
```
The values are the same. (AB)T = BTAT

**3)** Use matrix methods and r coding demonstrated in class to solve the following system of equations.
                                               3X   - 5Y   =   18
                                                      4X   =   20
```{r}
R <- matrix(c(18,
              20), ncol = 1, byrow = TRUE)

 Q <- matrix(c(3, -5,
               4, 0), ncol = 2, byrow =TRUE)
 
 solve(Q)%*%R
```
X = 5, Y = -0.6

**4)** By definition,  the rank of a matrix is the maximum number of linearly independent column vectors. Find the rank of the following matrix.   Justify your answer

```{r}
library(Matrix)

 C <- matrix(c(1,  5,  8,  3, 
               2,  7,  2,  6,
               3,  1,  11, 9),
             ncol = 4, byrow = TRUE)  
 
rankMatrix(C)[1] 
```

The maximum number of linearly independent column vectors is 3.

Column vector 4 is a multiple of column vector 1 if column vector 1 is multiplied by 3. Hence, the columns of C are linearly dependent. Since there is just 1 column vector out of 4 column vectors that is linearly dependent, therefore, the maximum number of independent column vectors = 4 - 1 = 3. 


**5a)** Find the inverse of the following matrix

```{r}
A <- matrix(c(5, 1, 3,
              4, 0, 5,
              1, 9, 6),
            ncol = 3,
            byrow = TRUE)

solve(A)
```

**5b)** Check that the resulting matrix is indeed the inverse.

```{r}
round(solve(A) %*% A, digits = 0)
```
The inverse of a matrix multiplied by the matrix should give an Identity matrix as above.
