---
title: "Multple Regression Interaction Analysis"
author: "Johnson ODEJIDE"
date: "2023-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, include = TRUE)

library(tidyverse)
```

```{r}
df <- tribble(~"WorkExperience",   ~"LevelofEducation",     ~"AnnualIncome($Thousands)",
                  21,                       6,                       	 31.7,
        	       14,                      3,              		 17.9,
        	        4,                       8,              		 22.7,
                   16,                      8,              		 63.1,
                   12,                      4,               		33,
                   20,                      4,               		41.4,
                   25,                      1,               		20.7,
                   8,                        3,               		14.6,
                   24,                      12,             		 97.3,
                   28,                       9,              		72.1,
                  4,                         11,             		 49.1,
                  15,                       4,              		 52)

```

**1)**

Use and show R code to produce a full Multiple Regression Model. The response variable is Annual Income.  Also produce the summary table

```{r}
MLR.fit <- lm(`AnnualIncome($Thousands)` ~ WorkExperience + LevelofEducation, data = df)

summary(MLR.fit)
```


**2)**

Use and show R code to produce a full Multiple Regression Model that includes and interaction term. Also produce the summary table

```{r}
MLR.int <- lm(`AnnualIncome($Thousands)` ~ WorkExperience + LevelofEducation + WorkExperience:LevelofEducation, data = df)

summary(MLR.int)
```


**3)**

Using summary indicators from both models determine which model you would use to predict Annual Income.  Justify your answer.  Your analysis should have 6 to 7 sentences using and comparing significance of pvalues, R2, adusted R2, variable standard errors, residual standard errors and the overall significance of both models.

**Response**

I would use the model without interaction terms because WorkExperience and LevelofEducation are both significant at level of 0.05 with p-values of 0.0095 and 0.00055 respecitively. The coefficients are also positive. The R-squared are also high enough showing that 80% of the variation in the AnnualIncome can be explained by the model and 75% of the variation in the AnnualIncome can be explained by the significant variables in the model. Overall, the model is also significant at p-value < 0.005.
On the other hand, the second model is not recommended. Although in general, the model appears to be significant at p-value of 0.002. However, WorkExperience and LevelofEducation are not significant at the level of 0.05. Furthermore, the interaction term between WorkExperience and LevelofEducation is also not significant. Nonetheless, about 82.54% of the variability in the AnnualIncome can be explained by the model.
The residual standard errors and the variable errors of both models seem similar in a more general sense.