---
title: "Classwork - Multiple Linear Regression and Dummy Variable"
author: "Johnson ODEJIDE"
date: "2023-03-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)

library(tidyverse)
```

```{r}
dt<- tribble(
  ~"WorkExperience",   ~"LevelofEducation",    ~"AnnualIncome($Thousands)",
        21,                       6,                       	 31.7,
        14,                      3,              		 17.9,
        4,                       8,              		 22.7,
        16,                      8,              		 63.1,
        12,                      4,               		33,
        20,                      4,               		41.4,
        25,                      1,               		20.7,
        8,                        3,               		14.6,
        24,                      12,             		 97.3,
        28,                       9,              		72.1,
        4,                         11,             		 49.1,
        15,                       4,              		 52)
dt
```
## 1

For the Table given above:
Use and show R code to produce a full Multiple Regression Model. The response variable is Annual Income.  Also produce the summary table

```{r}
mlrm.fit <- lm(`AnnualIncome($Thousands)` ~ WorkExperience + LevelofEducation, data = dt)
summary(mlrm.fit)
```

## 2

Use and show R code to produce a full Multiple Regression Model that includes and interaction term. Also produce the summary table

```{r}
mlrm.int.fit <- lm(`AnnualIncome($Thousands)` ~ WorkExperience + LevelofEducation + (WorkExperience * LevelofEducation), data = dt)

summary(mlrm.int.fit)

# mlrm.int.fit2 <- lm(`AnnualIncome($Thousands)` ~ WorkExperience + LevelofEducation + WorkExperience:LevelofEducation, data = dt)

# summary(mlrm.int.fit2)
```


## 3

Using summary indicators from both models determine which model you would use to predict Annual Income.  Justify your answer.  Your analysis should have 6 to 7 sentences using and comparing significance of pvalues, R2, adusted R2, variable standard errors, residual standard errors and the overall significance of both models.

I would use the model without interaction terms because In the model without an interaction term, looking at the Multiple R-squared, about 80.26% of the variation in the Annual income can be explained by the model and the Adjusted R-squared also indicates that 75.87% of the variation in the Annual Income can be explained by the model using only the dependent variables that are significant. Work Experience has a positive effect and the effect is significant at p = 0.0095. Similarly, the Level of Education has a positive effect and the effect is significant at p-value = 0.000552. In contrast, Work Experience and Level of Education are not significant and so we cannot individual reject the null hypothesis even though about 82.54% of variation in the Annual Income can be explained by the model.

```{r}
anova(mlrm.fit, mlrm.int.fit)
```



