---
title: "Lab 02"
author: "Johnson ODEJIDE"
date: "2023-01-27"
output:
  word_document: default
  html_document: default
subtitle: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
library(dplyr)
```
## a.

### Calculation of the Confidence Inverval
se = 0.4690

df = 8

b +/- t(se)

t = 2.306004

4.0 +/-  2.306004(0.4690)

confidence interval = (4.0 - 2.306004(0.4690), 4.0 + 2.306004(0.4690))

= (2.9185, 5.0815)

```{r}
X = c(1, 0, 2, 0, 3, 1, 0, 1, 2, 0)
Y = c(16, 9, 17, 12, 22, 13, 8, 15, 19, 11)

d.data <- tibble(X, Y)

conf.95 <- qt(p=.025, df=8, lower.tail = FALSE)

# conf.95

lm.fit <- lm(Y ~ X, data = d.data)
# lm.fit

# summary(lm.fit)

# qt(p=.025, df=8, lower.tail = FALSE)
# 4.00 +/- 2.306(0.469)

upper_bound <- 4.00 + 2.306 * 0.469
lower_bound <- 4.00 - 2.306 * 0.469

conf.int <- c(lower_bound, upper_bound)
conf.int

```


## b.

>While the prediction interval predicts in what range a future observation will fall, the confidence interval shows in what range of values the prediction falls based on some data provided already. In summary, confidence interval predicts what is available within the limits of the data while prediction interval is able to predict the future.


## c.

```{r}
new_df <- data.frame(X = 19)

predict(object = lm.fit, newdata = new_df, interval = "prediction") %>% 
cbind(new_df)

predict(object = lm.fit, newdata = new_df, interval = "confidence") %>% 
cbind(new_df)
```
***Prediction Interval when X = 19***

(66.4033, 105.9967)

***Confidence Interval when X = 19***

(66.701, 105.699)
