---
title: "Lab Model Building"
author: "Johnson Odejide"
date: "2023-04-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = T, message = F, warning = F)
library(tidyverse)
```

## 1

Using the diamonds data table, create a data table having only the variables `price`, `carat`, `x`, `y`, and `z`

```{r}
diamonds.dt <- diamonds[, c("price", "carat", "x", "y", "z")]
```

# 2 

Produce a multiple linear regression model using x, y, and z to predict price.

```{r}
fit.mlr <- lm(price ~ x + y + z, data = diamonds.dt)
summary(fit.mlr)
```

# 3 

Produce a multiple linear regression model using carat, x, and z to predict price.

```{r}
fit.mlr.c <- lm(price ~ carat + x + z, data = diamonds.dt)
summary(fit.mlr.c)
```

# 4 

Produce summary tables for the models that you generated for # 2 and # 3. Determine which does a better predicting price.  Justify your answer.

**Answer**

Both models have all predictors as significant. However, the second model (#3) that includes `carat`, `x`, `y`, and `z` as the predictors is better because it has the higher multiple R-squared and adjusted r-squared of 0.8541 and 0.8541 respectively compared to the other model that has 0.7825 and 0.7825 respectively. Both models are significant.